\chapter{Creating Your Source File}
\markright{Creating Your Source File} % new right header
%======================================================================
\section{Description of Source File(s)}
%======================================================================
\LaTeX\ source files are just ASCII text files containing the text of your document and the \LaTeX\ mark-up tags.

The mark-up tags are of two types, \emph{commands} and \emph{environments}.
\begin{description}
\item[Command] A command is a single \LaTeX\ tag which performs a specific action, such as inserting some blank space, or toggles a formatting mode on or off, such as a larger type face.
\item[Environment] An environment is a formatting mode which is delimited by \cmmd{begin} and \cmmd{end} statements.
For example, the \enviro{description} environment formats the display of a description offset from the surrounding text. 
Most commands also have an equivalent environment form.
\end{description}
%======================================================================
\section{Organizing Your Source File}
%======================================================================
Some general notes about organization:
\begin{itemize}
\item Clearly denote and comment the sections of your source file.
\item In the main body, put each sentence on a new line, and let the lines wrap.
This makes it easier to edit and move sentences around.
\item Plan your document ahead of time.
Think about what nomenclature you will use.
What \emph{logical structures} are suggested by the nomenclature? 
\item Will you need a bibliography? Do you have a lot of references which you will use again sometime?
\item Will you need an index? If so, plan concepts and words to index.
\end{itemize}
\LaTeX\ can easily keep track of all of these things, but it's easier to create your document if your are thinking about index items, citations and cross-references as you go, rather than as an afterthought! 
%======================================================================
\section{Editing Your Source File}
%======================================================================
Use your favourite text editor.
\begin{itemize}
\item Unix: \program{vi}, \program{emacs}, or the easy \program{pico}
\item PC or Mac: use any editor or word processor (and save as text)
\end{itemize}
Break your document up into several source files, a ``master'' file, and ``subdocuments'' (\eg\ one for each chapter and appendix).
%======================================================================
\section{Parts of the Master Source File}
%======================================================================
\subsection{Preamble}
%----------------------------------------------------------------------
The ``preamble'' of the source file tells the \program{latex} program how to format the structures and commands it finds in your document.

The basic format of the document is provided by the first preamble command, the \cmmd{documentclass} command.
Document classes built-in to \LaTeX\ are \textbf{letter, article, report, book,} and \textbf{slides}.
% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsubsection{Examples of the \cmmd{documentclass} Command}
% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
% the \verbatim environment displays the text (i.e. not in-line)
\begin{verbatim}
\documentclass[12pt,2col]{article}
\documentclass[10pt]{report}
\end{verbatim}
Options to this command (and other \LaTeX\ commands) are given in square brackets just before the curly braces.
Options are used to adjust the default point size of text, or other characteristics defined by a class.

UW has added its own modifications to some classes, notably, the \textbf{thesis91e} class.
However, this class uses a several ``styles'' dating from \LaTeX 2.09, which are not now available on many Unix hosts.  
Since all \textbf{thesis91e} does is prepare the front material for the thesis,
which is easily done ``by hand'', it is not necessary.
This document was prepared with the standard \textbf{report} document class.

The formatting commands may be built-in, defined in the standard packages, or defined by the user in the preamble itself with \cmmd{newcommand}. 
The \cmmd{usepackage} command identifies standard \LaTeXe\ packages as well as non-standard 
\footnote{
Note that all packages and options are plain (ASCII) text files written in \LaTeX\ or \TeX , which can be copied and modified by the user.
If you create your own packages or classes, \LaTeX\ will find them if you put them in the same directory as your document source files (or, under Unix, set \program{latex}'s path variables to find them).
} % end footnote
packages (which you could write yourself).
Don't forget to provide non-standard packages along with the other \LaTeX\ source documents if you give them to someone else!
%----------------------------------------------------------------------
\subsection{The Logical Document}
%----------------------------------------------------------------------
The logical document resides inside the \enviro{document} environment.
The document consists, usually, of three parts:
\begin{enumerate}
\item Front material --- title page, table of contents, list of figures, \etc
\item Main body of text --- chapters, sections, subsections, \etc
\item End material --- appendices, bibliography, index
\end{enumerate}
The outline of the logical document should be clear in the master source file.
Use the \cmmd{include} command, \eg\ 
\verb=\include{chap1}=, to include source subdocument \texttt{chap1.tex}.
% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsubsection{Example Master Source File}
% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\label{ex.master}
% paste in course.tex here
\input{mastersource}
